Seismic Time Series Analysis\\Statistics for Spatio-Temporal Data\\\textit{-- project draft --}

### 2020-06-09 draft.tex
more fixes on the draft and project code

Inspired by an article published during the initial phases of the Belgium-wide lockdown \cite{NatureCoronavirusSeismic}, we want to assess whether the lockdown period can be correlated with a visible change in seismic activity.
It is important to notice that some industrial events, such as mine explosions, are already mapped by seismologic institutions since they interfere with seismic surveys \cite{OtherSeismicEvents}. Such events are sometimes comparable with natural seismic events, depending on the distance from the seismograph.

@@

The vertical ground displacement is represented by the minimum and maximum displacements of the ground in the vertical axis, with $nm$ as unit of measure. This can easily be transformed to total ground displacement for that second by their difference.
For each day, a pair per second is available (86400 per day, around 30 million per year).

Out of the 9 stations, the following 6 have interesting positions, related to possible correlations with human activity:
\begin{itemize}[topsep=0.5em,itemsep=0em,partopsep=0.5em]
	\item \textbf{Uccle}, \textbf{Sart-Tilman}, and \textbf{Ében-Émael} are near a city (Bruxelles, Liege, and Maastricht).
	\item \textbf{Ostenda} This station is near the city of Ostenda, that borders with the sea.
	\item \textbf{Membach} and \textbf{Dourbes} are near a natural park.
\end{itemize}

@@

It is important to notice that the daily data in each file refers to the UTC time instead of the local Europe/Brussels time, an optimal choice to publish data in a machine-readable way, even if this could cause some issues while interpreting the results.

\subsection{Data cleanup and transformation}\label{sec:cleanup-transform}
\subsubsection{Data gathering and cleanup}
The data is provided in CSV format, one file per day, with only two columns related to the min-max pair \cite{Data20200402} and no column related to time. This causes two problems:
\begin{enumerate}[topsep=0.5em,itemsep=0em,partopsep=0.5em]
	\item Before merging different days, we need to add a column with the complete datetime for each row
	\item Days with some \texttt{NA} values cannot skip rows, so we will find placeholder values that indicate them $(\min, \max) \in \{(0,0), (-1,1)\}$
\end{enumerate}

Data from each station have different domains, as distinct equipment can register a specific amplitude of the seismic movements. In fact, we found some values outside the domains, transformed in \texttt{NA}.

In particular, we find that out of $\approx50$ million seconds, from 27-10-2018 to 28-05-2020, only 0.67\% (Uccle) and 0.37\% (Membach) have invalid values. We will see later how they are distributed for each station.

@@

\subsubsection{Data aggregation and initial transformation}
Since we are not interested in a second-precise analysis, and it would be computationally expensive to deal with the whole dataset, we aggregate data by hour. The aggregation will result in around 14 thousand points, that can still be captured by interpretable models.

While performing this operation, we should consider the correct time-zone: any seasonal analysis dealing with UTC data would be out of phase between the half-year when Belgium follows the UTC+01 timezone, and the half-year when it follows the UTC+02 timezone. With the help of the \texttt{lubridate} R package we converted to the actual timezones, and then we kept the actual objects without being timezone explicit, because all the methods we use are smart enough to work on the UTC conversion of datetimes. This transformation made it possible to have, for example, the time 6:00 in the timeseries to represent the local time in Belgium (\textit{Europe/Bruxelles}), without being influenced by the actual time-zone on that specific day.

The aggregation was performed considering the mean ground displacement, for the following reasons:
\begin{itemize}[topsep=0.5em,itemsep=0em,partopsep=0.5em]
	\item When an hour contains some missing seconds, these do not heavily affect the aggregation result, in contrast with summing the differences
	\item Seismic events, such as quakes and mine explosions, normally last some minutes at most, and will affect the result only slightly, as opposed to the hourly maximum
	\item Usually, maximum value aggregation does not follow the Normal distribution, while some models that we are going to use assume the data follow a Normal distribution
\end{itemize}

@@

In particular, after a visual analysis of the decomposition results, we can suppose the 90 hours gap from the Uccle station is related to some changes to the seismic equipment, that resulted in slightly more precise measurements.

\begin{figure}[h]
	\begin{subfigure}{.5\linewidth}
		\includegraphics[width=\linewidth]{project.uccs_files/figure-latex/missing values hour analysis-1.pdf}
		\caption{Uccle station}
		\label{missing-values-seconds:uccs}
	\end{subfigure}
	\begin{subfigure}{.5\linewidth}
		\includegraphics[width=\linewidth]{project.mems_files/figure-latex/missing values hour analysis-1.pdf}
		\caption{Membach station}
		\label{missing-values-seconds:mems}
	\end{subfigure}
	\caption{Missing values representation}
	\label{missing-values-seconds}
\end{figure}

\subsubsection{Seasonality assumptions}
The already cited article \cite{NatureCoronavirusSeismic} suggests the existence of a weekly seasonality. Looking at the data it seems confirmed, in addition to a daily seasonality.

Since our main focus is on the seasonalities themselves, we aim to prevent assumptions based on this objective. In other words, we want to avoid searching for correlation with human activity while imposing human-based seasonality periods. Thus, we will use an analytical method to extract seasonalities that the data itself suggests to us. Still, we recognize that the already performed transformations impose an human-based time alteration to the data itself (Section \ref{sec:cleanup-transform}).

The method we are going to apply is based on the Fourier Transform, and generates a \textit{periodgram} highlighting the most interesting frequencies that might correspond to a seasonality component.
Applying the FT for any frequency, we transform our 1-dimensional data into a 2-dimensional data mapped around a circle. Then, we can sum all the resulting vectors (that are our new data points) to obtain a vector for each frequency, whose absolute value represents the power level of the same frequency. An high power level represents a significant periodicity on that frequency in our original data.

A common operation to increase the difference between significant and non-significant periodicities is to apply a MA smoothing filter with a smaller window than the interesting ones. We will use such a filter with $f=2$.

@@

% From the Membach station, we can see an anomalous spike at a 6-hour periodicity.

@@

This initial division demostrated to be computationally heavy while fitting some models with weekly seasonality, and we noticed a possible yearly seasonality on Uccle data (\figurename~\ref{fig:whole-data-decomposition}). Even if it is not demonstrable given the short period our data covers, applying a 24-hour MA smoothing filter on the whole dataset (546 days) and then producing the periodgram as previously done, it suggests a 8748-hours significant periodicity (364.5 days).

\begin{figure}[h]
	\centering
	\includegraphics[width=0.7\linewidth]{"project.uccs_files/figure-latex/whole data plot-2"}
	\caption{Whole Uccle data decomposition}{\footnotesize Weeks 10 and 62 include January $1^{st}~2019/2020$}
	\label{fig:whole-data-decomposition}
\end{figure}

Given those issues, we choose the training/testing sets to hold 52/18 weeks: this is the greater train-to-test ratio we deem usable ($\approx 75\%/25\%$), while maintaining computational feasibility, and we hope it will render the possible yearly seasonality less intrusive.

@@

In \figurename~\ref{fig:direct-multi-seasonality} we see both approaches applied to both stations. Those plots show an interesting property of our seasonalities:
\begin{itemize}[topsep=0.5em,itemsep=0em,partopsep=0.5em]
	\item Both ways of extraction result in the same daily seasonality extracted
	\item The sum of the daily and weekly seasonalities extracted together is the same as the weekly seasonality extracted directly
\end{itemize}
Given in particular the second property, we will only use the 168-hours seasonality while fitting our models, for both the stations.

In \figurename~\ref{fig:decomposed} we see the double decomposition results. While data from Uccle station have a clear significativity of both trend and seasonalities, data from Membach could seem not significant. Still, if we notice that the main timeseries lies between values of 0 and 100, we can find out that around half of the value will be composed by the trend, a 30\% by the joined seasonality components, and the remaining value by the residuals. We will need to better handle the spikes that start from the middle of that dataset.

\section{Methodology}
Having two different datasets, we need to fit two separate models that represent the two stations. During this estimation, our aim is to obtain a model that generalises our training dataset in such a way that forecasting becomes effective, also assessing errors analytically.
Obtaining such a reliable model will be the first step to verify any significant difference forecasting over the lockdown period.

The first model we are going to estimate is the ARIMA model, including the seasonality components. To identify the order of the different components described by ARIMA, we will analyse the autocorrelation and partial autocorrelation of our datasets.
As a reminder:
\begin{itemize}[topsep=0.5em,itemsep=0em,partopsep=0.5em]
	\item non-seasonal AR process of order $p$ is described by an exponentially or sigmoidally decaying ACF, and by a cut-off of significant PACF values at lag $p$
	\item non-seasonal MA process of order $q$ is described by an exponentially or sigmoidally decaying PACF, and by a cut-off of significant ACF values at lag $q$
	\item seasonal AR process of order $P$ is described by an exponentially or sigmoidally decay at seasonal lags on ACF values, and by a cut-off of significant PACF values at each seasonal lag until the $P^{th}$
	\item seasonal MA process of order $Q$ is described by an exponentially or sigmoidally decay at seasonal lags on PACF values, and by a cut-off of significant ACF values at each seasonal lag until the $Q^{th}$
	\item given how significance is assessed for the ACF/PACF spikes, one over 20 spikes (5\%) could be \textit{coincidentally significant} or \textit{coincidentally \textbf{non}-significant}
\end{itemize}

@@

Since our dataset shows a relevant trend component, resulting in non-constant mean, we consider the use of the \textit{diff}\footnote{The \textit{diff} operator applies the \textit{d}-order differences, by default on lag 1. All \textit{diff} operators we will apply use $d=1$.} operator to smooth the mean.
After applying the \textit{diff} operator with lag 168, in conjunction with the lag 1 \textit{diff} or not, we come up with two possible models by checking the ACF and PACF plots:
\begin{itemize}[topsep=0.5em,itemsep=0em,partopsep=0.5em]
	\item from using both \textit{diff}s, the ACF and PACF on the long run hint to $\mathrm{ARIMA}(0,1,0)(0,1,1)_{168}$
	\item from using just the lag 168 \textit{diff}, the PACF on the first lags and the ACF/PACF combination on the long run hint to $\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$
\end{itemize}
We also employ the \textit{auto.arima} stepwise methodology\footnote{The \textit{auto.arima} stepwise methodology fits different ARIMA models within a range of orders, and based on each result's information criteria it will try to minimize it by changing each time one of the \textit{p}, \textit{d}, \textit{q}, \textit{P}, \textit{D}, or \textit{Q} values, depending on the configured limits.}, locking the \textit{diff} values to 0 or 1, to get some hints on the possible models.

Following both the analytical ways and the \textit{auto.arima} hints, and refining the models through further inspection of ACF/PACF plots of the residuals, we end up with three characteristic models. We selected them through the Akaike Information Criterion (AIC), being the only three out of the fitted models with $\mathrm{AIC} < 120.000$: $\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$, $\mathrm{ARIMA}(3,1,1)(0,1,1)_{168}$, and $\mathrm{ARIMA}(4,1,2)(0,1,1)_{168}$.

@@

In this case, the employment of \textit{auto.arima} stepwise methodology does not lead to any useful model, since it obtains $\mathrm{AICc}=\inf$ for the two manually-chosen models, that have an AICc value around -84.500, while the best model found by that methodology has an AICc around -81.000. This can be caused by approximation errors\footnote{The \textit{auto.arima} methodology was only used with initial approximation, due to high computing time.}, as we supposed earlier.

>>>>

Inspired by an article published during the initial phases of the Belgium-wide lockdown \cite{NatureCoronavirusSeismic}, we want to assess whether the lockdown period can be correlated with a visible change in background seismic activity.

It is important to notice that some industrial events, such as mine explosions, are already mapped by seismologic institutions since they interfere with seismic surveys \cite{OtherSeismicEvents}. Such events are sometimes comparable with natural seismic events, depending on the distance from the seismograph.

@@

The vertical ground displacement is represented by the minimum and maximum displacements of the ground in the vertical axis registered in a second, with $nm$ as unit of measure. This can easily be transformed to total ground displacement for each second by taking their difference. There are 86400 measurements per day, around 30 million per year.

Out of the 9 stations, the following 6 have interesting positions, related to possible correlations with human activity:
\begin{itemize}[topsep=0.5em,itemsep=0em,partopsep=0.5em]
	\item \textbf{Uccle}, \textbf{Sart-Tilman}, and \textbf{Ében-Émael} are near a city (Brussels, Liege, and Maastricht).
	\item \textbf{Ostenda} This station is near the city of Ostenda, that borders with the sea.
	\item \textbf{Membach} and \textbf{Dourbes} are near a natural park.
\end{itemize}

@@

It is important to notice that the daily data in each file refers to the UTC time instead of the local Europe/Brussels time, an optimal choice to publish data in a machine-readable way. This causes an issue fitting a fixed seasonality, when it depends on the legal time rather than on the absolute time.

\subsection{Data cleanup and transformation}\label{sec:cleanup-transform}
\subsubsection{Data gathering and cleanup}
The data is provided in CSV format, one file per day, with only two columns related to the min-max pair \cite{Data20200402} and no column related to time. This causes two problems:
\begin{enumerate}[topsep=0.5em,itemsep=0em,partopsep=0.5em]
	\item For merging days, we need to add a column with the second-precise datetime
	\item In days with missing values there are placeholder values instead a missing row, that would make the following rows misinterpreted for measurements on the wrong second. The placeholder values are one of these two min-max pairs: $(0,0)\, (-1,1)$
\end{enumerate}

Data from each station have different domains, as distinct equipment can register a specific amplitude of the seismic movements. In fact, we found some values\footnote{Values out of domain for both stations are less than 10 out of $\approx50$ million.} outside the domains, and we considered them as missing values.

In particular, we find that out of $\approx50$ million seconds, from 27-10-2018 to 28-05-2020, only 0.67\% (Uccle) and 0.37\% (Membach) have missing values. We will see later how they are distributed for each station.

@@

\subsubsection{Data aggregation and initial transformation}
Since we are not interested in a second-precise analysis, and it would be computationally expensive to deal with the whole dataset, we aggregate data by hour. The aggregation results in around 14 thousand points, that can still be captured by interpretable models.

While performing this operation, we should consider the correct time-zone: any seasonal analysis dealing with UTC dates would be out of phase for the half-year of legal time, with respect to the half-year of solar time. With the help of the \texttt{lubridate} R package we converted times to local time-zones, and then we kept them as absolute values, because all the R functions we use work on the UTC conversion of datetimes by default. This transformation made it possible to have each hour in the timeseries to represent the local time in Belgium (\textit{Europe/Brussels}), without being influenced by the legal time.

The aggregation was performed considering the mean total ground displacement, for the following reasons:
\begin{itemize}[topsep=0.5em,itemsep=0em,partopsep=0.5em]
	\item When an hour contains some missing seconds, these do not heavily affect the aggregation result, in contrast with summing the values
	\item Seismic events, such as quakes and mine explosions, normally last some minutes at most, and will affect the result only slightly, as opposed to the hourly maximum
	\item Usually, maximum value aggregation does not follow the Normal distribution, while some models that we are going to use assume the data follow a Normal distribution
\end{itemize}

@@

\begin{figure}[t]
	\begin{subfigure}{.5\linewidth}
		\includegraphics[width=\linewidth]{project.uccs_files/figure-latex/missing values hour analysis-1.pdf}
		\caption{Uccle station}
		\label{missing-values-seconds:uccs}
	\end{subfigure}
	\begin{subfigure}{.5\linewidth}
		\includegraphics[width=\linewidth]{project.mems_files/figure-latex/missing values hour analysis-1.pdf}
		\caption{Membach station}
		\label{missing-values-seconds:mems}
	\end{subfigure}
	\caption{Missing values representation}
	\label{missing-values-seconds}
\end{figure}

\subsubsection{Seasonality assumptions}
The already cited article \cite{NatureCoronavirusSeismic} suggests the existence of a weekly seasonality. Looking at the data it seems confirmed, in addition to a daily seasonality (\figurename~\ref{fig:zoom}).

Since our main focus is on the seasonalities themselves, we aim to prevent assumptions based on this objective. In other words, we want to avoid searching for correlation with human activity while imposing human-based seasonality periods. Thus, we use an analytical method to extract seasonalities that the data itself suggests to us. Still, we recognize that the already performed transformations impose an human-based time alteration to the data itself (Section \ref{sec:cleanup-transform}).

\begin{figure}[h]
	\begin{subfigure}{.5\linewidth}
		\includegraphics[width=\linewidth]{"project.uccs_files/figure-latex/whole data plot-2"}
		\caption{Uccle station}
		\label{fig:zoom:uccs}
	\end{subfigure}
	\begin{subfigure}{.5\linewidth}
		\includegraphics[width=\linewidth]{"project.mems_files/figure-latex/whole data plot-2"}
		\caption{Membach station}
		\label{fig:zoom:mems}
	\end{subfigure}
	\caption{Zoom from January $6^{th}$ to February $3^{rd}$ 2019}
	\label{fig:zoom}
\end{figure}

The method we are going to apply is based on the Fourier Transform (FT), and generates a \textit{periodgram} highlighting the most interesting frequencies that might correspond to a seasonality component.
Applying the FT for any frequency, we transform our 1-dimensional data into a 2-dimensional data mapped around a circle. Then, we can sum all the resulting vectors (that are our new data points) to obtain a vector for each frequency, whose absolute value represents the power level of the same frequency. An high power level represents a significant periodicity on that frequency in our original data.

A common operation to increase the difference between significant and non-significant periodicities is to apply a MA smoothing filter with a smaller window than the interesting ones. We will use such a filter with $p=2$.

@@

@@

Having the test set span from before to after the lockdown enables us to assess significant changes in the forecast performance on the two different periods.

This initial division demostrated to be computationally heavy while fitting some models with weekly seasonality, so that it would be challenging to fit more than a couple of models.
Given this issue, we choose the training/testing sets to hold 52/18 weeks: this is the greater train-to-test ratio we deem usable ($\approx 75\%/25\%$), while maintaining computational feasibility.
The assessing of the lockdown influence will be discussed in Section~\ref{sec:methodology}.

@@

In \figurename~\ref{fig:direct-multi-seasonality} we see both approaches applied to both stations. Those plots show interesting properties of our seasonalities, probably given by the method (local mean) and the fact that the weekly period is an exact multiple of the daily period:
\begin{itemize}[topsep=0.5em,itemsep=0em,partopsep=0.5em]
	\item Both ways of extraction result in the same daily seasonality extracted
	\item The sum of the daily and weekly seasonalities extracted together is the same as the weekly seasonality extracted directly
\end{itemize}
Given in particular the second property, we will only use the 168-hours seasonality while fitting our models, for both the stations.

In \figurename~\ref{fig:decomposed} we see the double decomposition results. While data from Uccle station have a clear significativity of both trend and seasonalities, data from Membach could seem not significant. Still, if we notice that the main timeseries lies between values of 0 and 100, we can find out that around half of the value will be composed by the trend, a 30\% by the joined seasonality components, and the remaining value by the residuals. We will need to apply some transformation to flatten the spikes that start from the middle of that dataset.

\section{Methodology}\label{sec:methodology}
Having two different datasets, we will operate on them separately and then we will compare the model interpretation.

For each station, we fit a model that generalises our training dataset in such a way that forecasting becomes effective, also assessing errors analytically.
Obtaining such a reliable model will be the first step to verify any significant difference forecasting over the lockdown period.

The model we are going to estimate is the ARIMA model, including the seasonality components. To identify the order of the different components described by ARIMA, we will analyse the autocorrelation and partial autocorrelation of our datasets.
As a reminder:
\begin{itemize}[topsep=0.5em,itemsep=0em,partopsep=0.5em]
	\item non-seasonal AR process of order $p$ is described by an exponentially or sigmoidally decaying ACF, and by a cut-off of significant PACF values at lag $p$
	\item non-seasonal MA process of order $q$ is described by an exponentially or sigmoidally decaying PACF, and by a cut-off of significant ACF values at lag $q$
	\item seasonal AR process of order $P$ is described by an exponentially or sigmoidally decay at seasonal lags on ACF values, and by a cut-off of significant PACF values at each seasonal lag until the $P^{th}$
	\item seasonal MA process of order $Q$ is described by an exponentially or sigmoidally decay at seasonal lags on PACF values, and by a cut-off of significant ACF values at each seasonal lag until the $Q^{th}$
	\item given how significance is assessed for the ACF/PACF spikes, one over 20 spikes (5\%) could be \textit{coincidentally significant} or \textit{coincidentally \textbf{non}-significant}
\end{itemize}

@@

Since our dataset shows a relevant trend component, resulting in non-constant mean, we consider the use of the \textit{diff}\footnote{The \textit{diff} operator applies the \textit{d}-order differences. All \textit{diff} operators we apply use $d=1$.} operator to smooth the mean.
After applying the \textit{diff} operator with lag 168, in conjunction with the lag 1 \textit{diff} or not, we come up with two possible models by checking the ACF and PACF plots:
\begin{itemize}[topsep=0.5em,itemsep=0em,partopsep=0.5em]
	\item from using both \textit{diff}s, the ACF and PACF on the long run hint to $\mathrm{ARIMA}(0,1,0)(0,1,1)_{168}$
	\item from using just the lag 168 \textit{diff}, the PACF on the first lags and the ACF/PACF combination on the long run hint to $\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$
\end{itemize}
We also employ the \textit{auto.arima} stepwise methodology\footnote{The \textit{auto.arima} stepwise methodology fits different ARIMA models within a range of orders. Based on each result's information criterion (Akaike), it changes each time one of the \textit{p}, \textit{d}, \textit{q}, \textit{P}, \textit{D}, or \textit{Q} values, depending on the configured limits. These changes minimizes the information criteria of the final model.}, locking the \textit{diff} values to 0 or 1, to get some hints on the possible models.

Following both the analytical ways and the \textit{auto.arima} hints, and refining the models through further inspection of ACF/PACF plots of the residuals, we end up with three characteristic models. We selected them through the Akaike Information Criterion (AIC)\footnote{Actually, we could have used the corrected AIC as it is usually more precise, but the values of AIC and AICc are equal to the first decimal for all our models ($\lesssim 10^{-7}\%$ difference).}, being the only three out of the fitted models with $\mathrm{AIC} < 120.000$: $\mathrm{ARIMA}(2,0,0)(0,1,1)_{168}$, $\mathrm{ARIMA}(3,1,1)(0,1,1)_{168}$, and $\mathrm{ARIMA}(4,1,2)(0,1,1)_{168}$.

@@

In this case, the employment of \textit{auto.arima} stepwise methodology\footnote{The \textit{auto.arima} stepwise methodology fits different ARIMA models within a range of orders. Based on each result's information criterion (Akaike), it changes each time one of the \textit{p}, \textit{d}, \textit{q}, \textit{P}, \textit{D}, or \textit{Q} values, depending on the configured limits. These changes minimizes the information criteria of the final model.} does not lead to any useful model, since it obtains $\mathrm{AIC}=\inf$ for the two manually-chosen models, that have an AIC value around -84.500, while the best model found by that methodology has an AIC around -81.000. This can be caused by approximation errors\footnote{The \textit{auto.arima} methodology was only used with initial approximation, due to high computing time.}, as we supposed earlier.

### 2020-06-09 project.Rmd
more fixes on the draft and project code

block_start <- NULL
block_end <- NULL
blocks <- data.frame(start=integer(), end=integer())
for (i in 1:length(seis.missing.ids)) {
  if (is.null(block_start)) {
    block_start <- seis.missing.ids[i]
    block_end <- seis.missing.ids[i]
  } else if ((block_end + 1) == seis.missing.ids[i]) {
    block_end <- seis.missing.ids[i]
  } else {
    blocks[nrow(blocks) + 1, ] <- c(block_start, block_end)
    block_start <- seis.missing.ids[i]
    block_end <- seis.missing.ids[i]
  }
}
blocks[nrow(blocks) + 1, ] <- c(block_start, block_end)
blocks$interval <- seconds(blocks$end - blocks$start + 1)
blocks$start <- seis.first.second + seconds(blocks$start - 1)
blocks$end <- seis.first.second + seconds(blocks$end - 1)
plot(blocks$start, blocks$interval,
     col=ifelse(blocks$interval<3600, "darkgreen", "darkred"),
     main="Consecutive seconds of missing values",
     ylab="Seconds", xlab="Interval start",
     xaxt="n", xlim=c(min(seis$utc_second), max(seis$utc_second))
)
axis.POSIXct(1, at=seq.POSIXt(from=min(seis$utc_second), to=max(seis$utc_second), by = "1 month"), format="%m-%y", cex=0.8)
abline(h=3600, col="green", lty="dotted")
legend("topleft", legend=c("interval >= 1h", "interval < 1h", "1h"), col=c("darkred", "darkgreen", "green"), lty=c(NA,NA,3), pch=c(1,1,NA),cex=0.9, bty="n")
text(x=blocks$start[which(blocks$interval > 3600)], y=blocks$interval[which(blocks$interval > 3600)], labels = paste0(blocks$start[which(blocks$interval > 3600)], " - ", blocks$end[which(blocks$interval > 3600)]), cex=0.9, pos=4)
remove(seis.complete, seis.missing.ids)
```

@@

block_start <- NULL
block_end <- NULL
blocks <- data.frame(start=integer(), end=integer())
for (i in 1:length(seis.missing.ids)) {
  if (is.null(block_start)) {
    block_start <- seis.missing.ids[i]
    block_end <- seis.missing.ids[i]
  } else if ((block_end + 1) == seis.missing.ids[i]) {
    block_end <- seis.missing.ids[i]
  } else {
    blocks[nrow(blocks) + 1, ] <- c(block_start, block_end)
    block_start <- seis.missing.ids[i]
    block_end <- seis.missing.ids[i]
  }
}
blocks[nrow(blocks) + 1, ] <- c(block_start, block_end)
blocks$interval <- hours(blocks$end - blocks$start + 1)
blocks$start <- seis.first.second + seconds(blocks$start - 1)
blocks$end <- seis.first.second + seconds(blocks$end - 1)
plot(c(rbind(blocks$start, blocks$end, NA)), rep(hour(blocks$interval), each=3),
     col=ifelse(blocks$interval<3600, "darkgreen", "darkred"),
     main="Consecutive hours of missing values",
     ylab="Hours", xlab="Interval start (mm-'yy)",
     type="l", lwd=10,
     xaxt="n", xlim=c(min(seis$utc_second), max(seis$utc_second))
     )
axis.POSIXct(1, at=seq.POSIXt(from=min(seis$utc_second), to=max(seis$utc_second), by = "1 month"), format="%m-'%y", cex=0.8)
text(x=blocks$start, y=hour(blocks$interval), labels = paste0(format(blocks$start, "%y-%m-%d %H:%M")), cex=0.9, pos=2)
text(x=blocks$end, y=hour(blocks$interval), labels = paste0(format(blocks$end, "%y-%m-%d %H:%M")), cex=0.9, pos=4)
text(x=blocks$start + make_difftime(hour=sapply(hour(blocks$interval), "/", y=2)), y=hour(blocks$interval), labels = paste0(hour(blocks$interval), "H"), cex=0.9, pos=3)
text(x=blocks$start + make_difftime(hour=sapply(hour(blocks$interval), "/", y=2)), y=hour(blocks$interval), labels = paste0(hour(blocks$interval), "H"), cex=0.9, pos=1)
remove(seis.complete, seis.missing.ids)

@@

# Analysis
```{r whole data plot}
month.starts <- seq.POSIXt(from=floor_date(seis.h.start, 'month'), to=ceiling_date(seis.h.end, 'month'), by = "1 month")
hours.from.start <- as.numeric(month.starts-seis.h.start)*24+168
autoplot(seis.h.ts, type="l") +
  scale_x_continuous(breaks=hours.from.start/168, labels=format(month.starts, "%m-%y")) +
  ylab("mean hourly movement (nm)") + xlab(NULL) + ggtitle("Training set") + labs(caption = "Data in original scale")
dec.addi <- decompose(seasonality.aggregate(seis.h.ts))
autoplot(dec.addi) + labs(title="Decomposition of additive time series", subtitle = paste0("Seasonality ", seasonality.period)) + xlab("Weeks")
```

>>>>

block_start <- NULL
block_end <- NULL
blocks <- data.frame(start=integer(), end=integer())
for (i in 1:length(seis.missing.ids)) {
  if (is.null(block_start)) {
    block_start <- seis.missing.ids[i]
    block_end <- seis.missing.ids[i]
  } else if ((block_end + 1) == seis.missing.ids[i]) {
    block_end <- seis.missing.ids[i]
  } else {
    blocks[nrow(blocks) + 1, ] <- c(block_start, block_end)
    block_start <- seis.missing.ids[i]
    block_end <- seis.missing.ids[i]
  }
}
blocks[nrow(blocks) + 1, ] <- c(block_start, block_end)
blocks$interval <- seconds(blocks$end - blocks$start + 1)
blocks$start <- seis.first.second + seconds(blocks$start + 1)
blocks$end <- seis.first.second + seconds(blocks$end + 1)
plot(blocks$start, blocks$interval,
     col=ifelse(blocks$interval<3600, "darkgreen", "darkred"),
     main="Consecutive seconds of missing values",
     ylab="Seconds", xlab="Interval start (mm-'yy)",
     xaxt="n", xlim=c(min(seis$utc_second), max(seis$utc_second))
)
axis.POSIXct(1, at=seq.POSIXt(from=min(seis$utc_second), to=max(seis$utc_second), by = "1 month"), format="%m-'%y", cex=0.8)
abline(h=3600, col="green", lty="dotted")
legend("topleft", legend=c("interval >= 1h", "interval < 1h", "1h"), col=c("darkred", "darkgreen", "green"), lty=c(NA,NA,3), pch=c(1,1,NA),cex=0.9, bty="n")
text(x=blocks$start[which(blocks$interval > 3600)], y=blocks$interval[which(blocks$interval > 3600)], labels = paste0(blocks$start[which(blocks$interval > 3600)], " - ", blocks$end[which(blocks$interval > 3600)]), cex=0.9, pos=4)
remove(seis.complete, seis.missing.ids)
```

@@

block_start <- NULL
block_end <- NULL
blocks <- data.frame(start=integer(), end=integer())
for (i in 1:length(seis.missing.ids)) {
  if (is.null(block_start)) {
    block_start <- seis.missing.ids[i]
    block_end <- seis.missing.ids[i]
  } else if ((block_end + 1) == seis.missing.ids[i]) {
    block_end <- seis.missing.ids[i]
  } else {
    blocks[nrow(blocks) + 1, ] <- c(block_start, block_end)
    block_start <- seis.missing.ids[i]
    block_end <- seis.missing.ids[i]
  }
}
blocks[nrow(blocks) + 1, ] <- c(block_start, block_end)
blocks$interval <- hours(blocks$end - blocks$start + 1)
blocks$start <- seis.first.second + hours(blocks$start + 1)
blocks$end <- seis.first.second + hours(blocks$end + 1)
plot(c(rbind(blocks$start, blocks$end, NA)), rep(hour(blocks$interval), each=3),
     col=ifelse(blocks$interval<3600, "darkgreen", "darkred"),
     main="Consecutive hours of missing values",
     ylab="Hours", xlab="Interval start (mm-'yy)",
     type="l", lwd=10,
     xaxt="n", xlim=c(min(seis$utc_second), max(seis$utc_second))
     )
axis.POSIXct(1, at=seq.POSIXt(from=min(seis$utc_second), to=max(seis$utc_second), by = "1 month"), format="%m-'%y", cex=0.8)
text(x=blocks$start, y=hour(blocks$interval), labels = paste0(format(blocks$start, "%y-%m-%d %H:%M")), cex=0.9, pos=2)
text(x=blocks$end, y=hour(blocks$interval), labels = paste0(format(blocks$end, "%y-%m-%d %H:%M")), cex=0.9, pos=4)
text(x=blocks$start + make_difftime(hour=sapply(hour(blocks$interval), "/", y=2)), y=hour(blocks$interval), labels = paste0(hour(blocks$interval), "H"), cex=0.9, pos=3)
text(x=blocks$start + make_difftime(hour=sapply(hour(blocks$interval), "/", y=2)), y=hour(blocks$interval), labels = paste0(hour(blocks$interval), "H"), cex=0.9, pos=1)
remove(seis.complete, seis.missing.ids)

@@

# Analysis
```{r whole data plot}
month.starts <- seq.POSIXt(from=floor_date(seis.h.start, 'month'), to=ceiling_date(seis.h.end, 'month'), by = "1 month")
hours.from.start <- as.numeric(month.starts-seis.h.start)*24+168
autoplot(seis.h.ts, type="l") +
  scale_x_continuous(breaks=hours.from.start/168, labels=format(month.starts, "%m-%y")) +
  ylab("mean hourly movement (nm)") + xlab(NULL) + labs(title = "Training set", caption = "Data in original scale")
week.starts <- seq.POSIXt(from=floor_date(seis.h.start, 'week'), to=ceiling_date(seis.h.end, 'week'), by = "1 week")
hours.from.start <- as.numeric(week.starts-seis.h.start)*24+168
autoplot(seis.h.ts, type="l") +
  scale_x_continuous(breaks=hours.from.start/168, labels=format(week.starts, "%d-%m-%y"), limits = c(11,15)-3/24) +
  scale_y_continuous(limits = c(min(seis.h.ts[(10*168):(16*168)]), max(seis.h.ts[(10*168):(16*168)]))) +
  ylab("mean hourly movement (nm)") + xlab(NULL) + labs(title = "Training set", subtitle = "4-weeks zoom", caption = "Data in original scale")
dec.addi <- decompose(seasonality.aggregate(seis.h.ts))
autoplot(dec.addi) + labs(title="Decomposition of additive time series", subtitle = paste0("Seasonality ", seasonality.period)) + xlab("Weeks")
```

### 2020-06-09 project.after.Rmd
more fixes on the draft and project code

plot_grid(
  autoplot(ts(seasonbase.before.lockdown.24.df$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() + geom_line(aes(col="Direct daily"), size=1.5) +
    geom_line(aes(y=season.before.lockdown.24.df$figure, col="Multi daily")) +
    scale_colour_manual("", values=c("purple", "yellow"), breaks=c("Direct daily", "Multi daily")),
  autoplot(ts(seasonbase.before.lockdown.168.df$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() + geom_line(aes(col="Direct weekly"), size=1.5) +
    geom_line(aes(y=season.before.lockdown.168.df$figure + rep(season.before.lockdown.24.df$figure, 7), col="Multi w+d")) +
    #geom_line(aes(y=rep(seasonbase.before.lockdown.24$figure, 7), col="Direct daily"), size=1.5) +
    #geom_line(aes(y=rep(season.before.lockdown.24$figure, 7), col="Multi daily")) +
    geom_line(aes(y=season.before.lockdown.168.df$figure, col="Multi weekly")) +
    scale_colour_manual("", values=c("red", "purple", "blue", "yellow", "green"), breaks=c("Direct weekly", "Direct daily", "Multi" %s+% c(" weekly", " daily", " w+d"))),
  nrow = 2)

>>>>

plot_grid(
  autoplot(ts(seasonbase.before.lockdown.24.df$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() + geom_line(aes(col="Direct daily"), size=1.5) +
    geom_line(aes(y=season.before.lockdown.24$figure, col="Multi daily")) +
    scale_colour_manual("", values=c("purple", "yellow"), breaks=c("Direct daily", "Multi daily")),
  autoplot(ts(seasonbase.before.lockdown.168.df$figure, frequency=1)) + ylab("") + xlab("Hours") + scale_x_continuous() + geom_line(aes(col="Direct weekly"), size=1.5) +
    geom_line(aes(y=season.before.lockdown.168.df$figure + rep(season.before.lockdown.24$figure, 7), col="Multi w+d")) +
    #geom_line(aes(y=rep(seasonbase.before.lockdown.24$figure, 7), col="Direct daily"), size=1.5) +
    #geom_line(aes(y=rep(season.before.lockdown.24$figure, 7), col="Multi daily")) +
    geom_line(aes(y=season.before.lockdown.168.df$figure, col="Multi weekly")) +
    scale_colour_manual("", values=c("red", "purple", "blue", "yellow", "green"), breaks=c("Direct weekly", "Direct daily", "Multi" %s+% c(" weekly", " daily", " w+d"))),
  nrow = 2)
